---
title: "Create Clean Data"
author: "Nick chapman"
date: '(`r format(Sys.time(), "%B %d, %Y")`)'
output: 
  github_document: default
  html_notebook: default 
---

## Introduction
This notebook contains the code to download demographic data from `tidycensus` for the State of Missouri, a `.shp` file of the counties of Missouri from `tigris` and to clean the data from each of them.

## Dependencies
This notebook requires a few packages:

```{r load-packages}
# tidyverse packages
library(tidycensus)
library(tigris)
library(dplyr)
```

## Load Data
This notebook requires data downloaded, saved to the `data/raw/` folder, as well as data downloaded from `tigris` and `tidycensus`:

### Load Population Data

`B99187_001` is the variable for: “Estimate!!Total”, I will use this to get a Population for each county in Missouri, as well as a Population Density:
```{r load-data}
MO_population <- get_acs(geography = "county", state = "MO", year = 2015, variable = "B99187_001", survey = "acs5")
```

### Load Income Data

`B19326_001` is the variable for: “Median income in the past 12 months (in 2017 inflation-adjusted dollars)”:

```{r}
MO_Income <- get_acs(geography = "county", state = "MO", year = 2015, variable = "B19326_001", survey = "acs5")
```

