---
title: "Create Clean Data"
author: "Nick chapman"
date: '(`r format(Sys.time(), "%B %d, %Y")`)'
output: 
  github_document: default
  html_notebook: default 
---

## Introduction
This notebook contains the code to download demographic data from `tidycensus` for the State of Missouri, a `.shp` file of the counties of Missouri from `tigris`, read .csv files from Missouri's County Health Rankings, and Mophims and to clean the data from each of them.

## Dependencies
This notebook requires a few packages:

```{r load-packages}
# tidyverse packages
library(tidycensus)
library(tigris)
library(dplyr)
library(readr)
library(here)
library(naniar)
library(janitor)
library(sf)
```

## Load Data

This notebook requires data downloaded, saved to the `data/raw/` folder, as well as data downloaded from `tigris` and `tidycensus`:

### Load Spatial Data

I will load the data for the counties of Missouri, through `tigris`:

```{r load-Mo-Counties}
moCounties <- counties(state = 29, cb = FALSE)
```

Now, I need to convert it to an sf object:

```{r}
moCounties <- st_as_sf(moCounties)
```


### Load Population Data

`B99187_001` is the variable for: “Estimate!!Total”, I will use this to get a Population for each county in Missouri, as well as a Population Density:

```{r load-pop-data}
MO_population <- get_acs(geography = "county", state = "MO", year = 2015, variable = "B99187_001", survey = "acs5")
```

### Load Income Data

`B19326_001` is the variable for: “Median income in the past 12 months (in 2017 inflation-adjusted dollars)”:

```{r MO-County-Data}
MO_Income <- get_acs(geography = "county", state = "MO", year = 2015, variable = "B19326_001", survey = "acs5")
```

### Load Heart Attack Data

This heart attack data was downloaded from Mophim:

```{r read-Mophims-stats}
Heart_Attack_MO<- read_csv(here("data", "raw", "Heart_Attack_Raw_MOcounty.csv"))
```

### Load County Health Ranking Data

```{r read-county-ranking-csv}
county_rankings<- read_csv(here("data", "raw", "county_rankings.csv"))
```


## Cleaning Variables

I will first clean up my tables, removing unnecessary variables:

### Clean Heart Attack Data

For the heart attack data,`Mapped Value`, `X6`, and `Rank` are unnecessary, for I will rank them myself later:

```{r clean-heart-vars}
Heart_Attack_Clean<- select(Heart_Attack_MO, -"X6", -"Rank", -"Mapped Value" )
```

Now, I want to rename `FIPS Code` to `GEOID` and make it the same value as the income and population data:

```{r rename-heart-fips}
Heart_Attack_Clean <- rename(Heart_Attack_Clean, GEOID = "FIPS Code")
```

```{r transform-heart-geoid}
Heart_Attack_Clean <-  transform(Heart_Attack_Clean, GEOID = as.numeric(GEOID))
```

```{r mutate-heart-geoid}
Heart_Attack_Clean <- mutate(Heart_Attack_Clean, GEOID = GEOID +29000)
```

### Clean Income Data

For the Income data, `variable` and `moe` are unnecessary and I will remove them:

```{r select-income-vars}
MO_Income_Clean<- select(MO_Income, -"variable", -"moe")
```

### Clean Population Data

For the population data, `variable` is unnecessary and I will remove it:

```{r select-pop-vars}
MO_population_Clean<- select(MO_population, -"variable")
```

### Clean County Ranking Data

```{r select-county-health-vars}
County_Rankings_Clean <- select(county_rankings, "X1", "X3", "Adult smoking", "Adult obesity", "Food environment index", "Physical inactivity", "Access to exercise opportunities")
```

```{r rename-county-health}
County_Rankings_Clean <- rename(County_Rankings_Clean, GEOID = "X1")
County_Rankings_Clean <- rename(County_Rankings_Clean, County = "X3")
County_Rankings_Clean <- County_Rankings_Clean[-c(1,2),]
```


Now, all of my tables have the same 'GEOID', all of the county level data could now be joined together.

